---
title: "discrete_trait_analysis_summaries"
output: html_document
date: "2023-04-18"
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(kableExtra)

library(table1)
library(dplyr)
library(tidyr)

library(ggplot2)
library(ggpubr)
library(magrittr)

library(ape)
library(ggtree)
library(phangorn)

library(arsenal)
library(officer)

# Set file paths - if knitting, save in a folder by date
if (knitr::is_html_output()) {
  save_path <- paste0("../results/regional trends/", format(Sys.Date(), "%b%d_%Y"), "/")
} else {
  save_path <- "../results/regional trends/"
}
```
Info on phangorn package:   <https://academic.oup.com/bioinformatics/article/27/4/592/198887?login=false>  
<https://cran.r-project.org/web/packages/phangorn/phangorn.pdf>  
<br><br>



# Read in data
### metadata
```{r metadata}
data.hiv <- readRDS("../metadata/participants.rds") %>% filter(HIV_seq_available == "Yes" & !is.na(sampling_year.F)) # metadata for all sequences
data.SHARP_hiv <- readRDS("../metadata/participants.SHARP_hiv.rds") # metadata for SHARP HIV sequences
SHARP_seq_names <- data.SHARP_hiv$annotated_name
```
TO DO: find a clearer way to filter participant data to only seqs (ideally, this should be done in the metadata script)

### tree data
```{r tree data}
# SHARP only
tree.SHARP_hiv <- readRDS("../HIV/tree.SHARP_hiv.rds") #tree data object
tree.SHARP_hiv <- root(tree.SHARP_hiv, "D.2019.PWID.Nairobi.Male.OQ299352.SP222") #root at a random sequence for now
tree.SHARP_hiv.meta <- tree.SHARP_hiv.meta <- readRDS("../HIV/meta_tree.SHARP_hiv.rds") #metadata with info about tree 
# SHARP and published
tree.hiv <- readRDS("../HIV/tree.hiv.A1.rds") #tree data object
tree.hiv <- root(tree.hiv, "A1.2019.PWID.Nairobi.Male.OQ299131.SP001") #root at a random sequence for now  - did not work (why?) 
tree.hiv.meta <- readRDS("../HIV/meta_tree.hiv.A1.rds") #metadata with info about tree 
```  



<br><br><br>

# DATA SUMMARY


### Subtype by study
Counts of available seqs by study
```{r}
data.hiv %>% select(source, hiv_subtype.F) %>% table(useNA = "ifany") #USE ME

data.hiv %>% select(source, hiv_subtype2.F) %>% table(useNA = "ifany")
```

TO DO: check this  



### Table 1
```{r Table 1}
library("arsenal")

# Fix and addsome variables (should be in the data cleaning script)
#fix skip pattern for active injector variable
data.SHARP_hiv <- data.SHARP_hiv %>% mutate(
  activeinjector.F = case_when(everuseddrugs == 0  ~ "No",
                  TRUE ~ activeinjector.F) #People who don't use drugs are NA because of skip pattern, but should have been "No"
  )
#fix skip pattern for share_needles variable
data.SHARP_hiv <- data.SHARP_hiv %>% mutate(
  shareneedles.F = case_when(everuseddrugs == 0  ~ "have not injected in the last month",
                  TRUE ~ shareneedles.F) #People who don't use drugs are NA because of skip pattern, but should have been "No"
  )

#add a time since diagnosis variable that excludes those not previously diagnosed
data.SHARP_hiv <- data.SHARP_hiv  %>%
  mutate(years_since_HIV_diagnosis2 = case_when(
    hivtestbeforepositive == "No" ~ as.numeric(NA),
    TRUE ~ years_since_HIV_diagnosis
  )) #Make a variable for years since first diagnosis that is NA for anyone diangosed at enrolement 


# Table 1 controls

tableby_controls <- tableby.control(
    numeric.stats = c("Nmiss", "meansd", "medianq1q3"),
    cat.stats = c("Nmiss", "countpct"),
    digits = 1,
    digits.pct = 1)


# By region (SHARP)
#Seperate rcdmoneysex.MF varaible into 2 varaibles by sex (that way we can get seperate p-values)
data.SHARP_hiv <- data.SHARP_hiv %>%
  mutate(
    rcdmoneysex.M = case_when(
      is.na(rcdmoneysex.MF) ~ as.character(NA),
      rcdmoneysex.MF == "F_No" | rcdmoneysex.MF == "F_Yes" ~ as.character(NA),
      rcdmoneysex.MF == "M_No" ~ "No",
      rcdmoneysex.MF == "M_Yes" ~ "Yes"),
    rcdmoneysex.F = case_when(
      is.na(rcdmoneysex.MF) ~ as.character(NA),
      rcdmoneysex.MF == "M_No" | rcdmoneysex.MF == "M_Yes" ~ as.character(NA),
      rcdmoneysex.MF == "F_No" ~ "No",
      rcdmoneysex.MF == "F_Yes" ~ "Yes")
    )

table1_regional_trends.SHARP_hiv <- tableby(region.CN ~
    hiv_status_final +
    hivtestbeforepositive +
    years_since_HIV_diagnosis +
    years_since_HIV_diagnosis2 + 
    hcvresults_final +
    hiv_subtype.F +
    hiv_subtype2.F +
    sex +
    rcdmoneysex.M +
    rcdmoneysex.F +
    everinjecteddrugs +
    activeinjector.F +
    shareneedles.F +
    sex_partners_3mo.cat +
    risk2.F +
    risk4.F +
    barriertransportation + 
    stablehousing + 
    sleeppastmonth.F +
    sevennightsawaynairobi,
  data = data.SHARP_hiv,
  control = tableby_controls
)

summary(table1_regional_trends.SHARP_hiv)

setwd(save_path) #This only changes the working directory within the chunk - necessary because write2Word does not accept a file path
write2word(table1_regional_trends.SHARP_hiv, "region_SHARP_table_1.hiv.doc")




# By region (all)

table1_regional_trends.hiv <- tableby(region ~
    source +
    sampling_year.cat +
    risk2.F +
    risk4.F +
    region +
    hiv_subtype.F +
    hiv_subtype2.F +
    sex,
  data = data.hiv,
  control = tableby_controls
)

summary(table1_regional_trends.hiv)

write2word(table1_regional_trends.hiv, "region_table_1.hiv.doc")



# By region (Coast and Nairobi only)

table1_regional_trends.hiv.CN <- tableby(region ~
    source +
    sampling_year.cat +
    risk2.F +
    risk4.F +
    region +
    hiv_subtype.F +
    hiv_subtype2.F +
    sex,
  data = data.hiv[!is.na(data.hiv$region.CN),],
  control = tableby_controls
)

summary(table1_regional_trends.hiv.CN)

write2word(table1_regional_trends.hiv.CN, "region_table_1.hiv.CN.doc")




# By risk group (all)

#Table 1 - HIV
table1_risk_trends.hiv <- tableby(risk4.F ~
    source +
    sampling_year.cat + 
    region +
    hiv_subtype.F +
    hiv_subtype2.F +
    sex,
  data = data.hiv,
  control = tableby_controls
)

summary(table1_risk_trends.hiv)

write2word(summary(table1_risk_trends.hiv), "risk_table_1.hiv.doc")


# By risk group (coast and Nairobi only)

#Table 1 - HIV
table1_risk_trends.hiv.CN <- tableby(risk4.F ~
    source +
    sampling_year.cat + 
    region +
    hiv_subtype2.F +
    sex,
  data = data.hiv[!is.na(data.hiv$region.CN),],
  control = tableby_controls
)

summary(table1_risk_trends.hiv.CN)

write2word(summary(table1_risk_trends.hiv.CN), "risk_table_1.hiv.CN.doc")


# By sampling_year (all)

#Table 1 - HIV
table1_year_trends.hiv <- tableby(sampling_year.cat ~
    risk2.F +
    source +
    region +
    hiv_subtype.F +
    hiv_subtype2.F +
    sex,
  data = data.hiv,
  control = tableby_controls
)

summary(table1_year_trends.hiv)

write2word(table1_year_trends.hiv, "year_table_1.hiv.doc")


setwd("../../scripts/") #reset working directory just in case


```

TO DO: confirm what test is used to get p-value  
TO DO: generate "years_since_first_hiv_diagnosis2" in data cleaning script and fix everinjecteddrugs variable (to be no and not NA if they never used drugs) in the data cleaning script  

# Other summary
```{r}
# Confirm participants who are NA for ever injecting drugs are partners and not included in PWID data - correct
data.SHARP_hiv %>%
  filter(is.na(everinjecteddrugs.F)) %>%
  select(everuseddrugs, everinjecteddrugs, everinjecteddrugs.F, index_or_partner.F, partner_type_comb.F, partner_type, MSM.F, FSW.F, PWID.F, PWID_region.comb) %>% View()



# How many participants were known pos?
#Among all SHARP participants living with HIV
data.SHARP_hiv %>% select(hivtestbeforepositive, index_or_partner.F) %>% table(useNA = "ifany")
#AMong SHAPR participants with seqs
data.SHARP_hiv %>% select(hivtestbeforepositive, index_or_partner.F) %>% table(useNA = "ifany")



#Year range for each population group (published data)
data.hiv %>% filter((hiv_subtype.F == "A1" | hiv_subtype.F == "C" | hiv_subtype.F == "D") &
                      !is.na(region.CN) & source == "published") %>% group_by(risk4.F) %>%
  summarise(range = range(sampling_year.F), n = n())

data.hiv %>% filter((hiv_subtype.F == "A1" | hiv_subtype.F == "C" | hiv_subtype.F == "D") &
                      !is.na(region.CN) & source == "published") %>% select(sampling_year.cat) %>% table(useNA = "ifany")


#Distribution of population groups in SHARP data after removing recombinant sequences
data.hiv %>% filter((hiv_subtype.F == "A1" | hiv_subtype.F == "C" | hiv_subtype.F == "D") & source == "SHARP") %>%
  select(risk2.F) %>% table()

```


<br><br>


### Associations 
Not using - decided to just use the chi-sq test implemented in Arsenal (above)  
Note: all variables must be factors for this to work  
```{r Regression on region}
#factor barriertransportation variable
data.SHARP_hiv$barriertransportation.F = factor(data.SHARP_hiv$barriertransportation, levels = c("No", "Yes"))
#drop B subtype from data.SHARP_hiv because there are no B usbtypes in this data, so this causes an error
data.SHARP_hiv$hiv_COMET_subtype.F <- droplevels(data.SHARP_hiv$hiv_COMET_subtype.F)

# Implement regression and add to a table
traits <- c("hiv_COMET_subtype.F", "everinjecteddrugs.F", "shareneedles.F", "sex_partners_3mo.cat", "risk2.F", "barriertransportation.F")

regression.df <- data.frame()

for(variable in traits){
  print(variable)
  formula <- as.formula(paste0("region.CN ~ ", variable))
  variable2 <- data.SHARP_hiv[,variable, drop=TRUE]
  my_glm <- glm(formula, data = data.SHARP_hiv, family = "binomial")
  OR <- exp(coef(my_glm))[2:length(coef(my_glm))]
  CI_2.5 <- exp(confint(my_glm))[,1][2:length(coef(my_glm))]
  CI_97.5 <- exp(confint(my_glm))[,2][2:length(coef(my_glm))]
  signif <- ifelse((CI_2.5 > 1 & CI_97.5 > 1) | (CI_2.5 < 1 & CI_97.5 < 1), "*", "")
  effect <- paste0(round(OR,2), " (", round(CI_2.5, 2), "-", round(CI_97.5,2), ")", signif)
  
  if(!(is.factor(variable2) | is.numeric(variable2))){
    print("varaibles must be either numeric or factors")
    break()
  }
  
  if(is.factor(variable2)){
    total <- sum(!is.na(variable2))
    count <- table(variable2)
    percent <- paste0(round(100*count/total, 1), "%")
    proportion <- paste0(count, " (", percent, ")")
    
    count_Coast <- table(data.SHARP_hiv[,variable, drop = TRUE], data.SHARP_hiv$region.CN)[,2]
    count_Nairobi <- table(data.SHARP_hiv[,variable, drop = TRUE], data.SHARP_hiv$region.CN)[,1]
    odds_CoastvNairobi <- round(count_Coast/count_Nairobi, 2)
    odds <- paste0(count_Coast, "/", count_Nairobi, " = ", odds_CoastvNairobi)
    
    variable_name <- ifelse(is.null(label(variable2)), "?", label(variable2))
    
    regression.row <- data.frame(variable =  c(variable_name, rep("", length(levels(variable2))-1)), #Note: levels only works with factors
                                 values = levels(variable2),
                                 mean_or_proportion = proportion,
                                 odds_CoastvsNairobi = odds,
                                 OR = rep("-", length(levels(variable2))))
    effect <- append("Ref", effect)
    
  } else {
    mean <- round(mean(variable2, na.rm = TRUE), 2)
    
    regression.row <- data.frame(variable = label(variable2),
                                 values = "-",
                                 mean_or_proportion = mean,
                                 odds_CoastvsNairobi = "-",
                                 OR = "-")
  }
  
  regression.row$OR <- effect
  regression.df <- rbind(regression.df, regression.row)
}

write.csv(regression.df, "save_pathregression.csv")
```

To do: check warnings. I believe these are all indicating low counts in the regression. I will exclude these from the table after the fact but need to verify  
TO DO: remove ste tp factorize barriertransportation and region.CN variable - this should be done in the clean_metadata script


<br><br><br><br>

# TREE VISUALIZATION by region  

## SHARP only
```{r plot region on SHARP tree, fig.height=30, fig.width=10}

# find node between A1 and C subtypes in order to rotate tree
ggtree(tree.SHARP_hiv)  + geom_text(aes(label=node), color = "red", size = 2, hjust = 1) 



#re-extract metadata (because I want to add the sharing needles variable) (ideally, this would be done in the combine_tree_and_metadata cleaning script)
variables <- c("ptid", "Sequence.name", "annotated_name", "tree_seq_name", "source", "index_or_partner.F", "index_ptid", "index_ptids.F", "partner_type_comb.F", "sex", "age", "enrollment_date", "sampling_year.F", "region", "region.CN", "rcdmoneysex", "everinjecteddrugs", "hiv_status_final.F", "hcvresults_final", "hiv_subtype.F", "hiv_subtype2.F", "risk", "risk2.F", "risk3.F", "risk4.F", "PWID_region.comb", "HET.F", "MSM.F", "FSW.F", "PWID.F", "hiv_vl", "shipment", "shareneedles", "activeinjector.F", "REGA_subtype.F", "REGA_subtype2.F")
tree.SHARP_hiv.tibble <- as_tibble(tree.SHARP_hiv)
tree.SHARP_hiv.meta <- left_join(tree.SHARP_hiv.tibble, data.SHARP_hiv[, variables], by = c("label" = "annotated_name"))


#Add identified for whether a "possible FSW" and whether shared needles
tree.SHARP_hiv.meta <- tree.SHARP_hiv.meta %>%
  mutate(possibleFSW.label = ifelse (risk2.F == "FSW_PWID", "*", ""),
         MSM.label = ifelse(risk2.F == "MSM_PWID", " *", ""),
         HET.label = ifelse(risk2.F == "HET" | risk2.F == "MSM" | risk2.F == "FSW", "  *", ""), #Note: decided not to include
         share_needles.label = ifelse (shareneedles == "Yes", "   *", ""))


# By region and risk
tree.SHARP_hiv <- root.tree(tree.SHARP_hiv)

tree.SHARP_hiv.meta  <-tree.SHARP_hiv.meta %>%
  mutate(REGA_subtype2.F = case_when(
    is.na(REGA_subtype2.F) & !is.na(source) ~ as.factor("recombinant"),
    !is.na(source) ~ as.factor(REGA_subtype2.F),
    TRUE ~ as.factor(REGA_subtype2.F))) #I think the missing values were recombinants - needs checking what happened here

region.SHARP_hiv.tree <- ggtree(tree.SHARP_hiv) %>% rotate(343) %<+% tree.SHARP_hiv.meta +
  geom_tippoint(aes(color = region.CN, shape = REGA_subtype2.F), size = 2) +
  scale_color_manual(values = c("forestgreen",  "deepskyblue"), labels = c("Coast", "Nairobi"), name = "Region") +
  scale_shape_manual(values = c("AE" = 14, "A1" = 15, "A2" = 7, "C" = 16, "Clike" = 13, "D" = 17,
                                "G" = 18, "recombinant" =6, "recombinantCRF" = 5),
                     labels = c("AE" = "AE", "A1" = "A1", "A2" = "A2", "C" = "C", "Clike" = "C-like", "D" = "D",
                                "G" = "G", "recombinant" = "recombinant", "recombinantCRF" = "recombinant CRF"),
                      name = "Subtype") + 
  geom_tiplab(aes(label = possibleFSW.label), color = "red", size = 10, nudge_y = -1.1) +
  geom_tiplab(aes(label = MSM.label), color = "yellow4", size = 10, nudge_y = -1.1) +
  geom_tiplab(aes(label = share_needles.label), color = "black", size = 10, nudge_y = -1.1) +
  theme(legend.text=element_text(size=15), legend.title=element_text(size=20)) +
  geom_treescale() 
region.SHARP_hiv.tree
ggsave(region.SHARP_hiv.tree, file = paste0(save_path, "region.SHARP_hiv.png"), width = 10, height = 10.5, units = "in") #CAUTION: For some reason ggsave is failing to print any points of a non-solid share. Bypass this by savign directly from the plot sindow


# By region and subtype
geneotypeXregion.SHARP_hiv.tree <- ggtree(tree.SHARP_hiv) %<+% tree.SHARP_hiv.meta +
  geom_tippoint(aes(colour = region.CN, shape = region.CN, fill = REGA_subtype2.F), stroke = 1.3) + 
  geom_treescale() +
  scale_colour_manual(values = c("deepskyblue", 000000, "black")) +
  scale_shape_manual(values = c(22,21,24)) +
  #There is a bug for the interaction between fill and scale_shap_manual - this is a fix to make the legend work
  guides(fill = guide_legend(override.aes=list(shape=c(21))),
         shape = guide_legend(override.aes=list(fill = "forestgreen")))

geneotypeXregion.SHARP_hiv.tree
ggsave(geneotypeXregion.SHARP_hiv.tree, file = "save_pathgeneotypeXregion.SHARP_hiv.png", width = 7, height = 10.5, units = "in")
```

## SHARP and published seqs
```{r plot region on SHARP and published tree, fig.height=60, fig.width=10, eval = FALSE}
tree.hiv.meta$source <- ifelse(is.na(tree.hiv.meta$source), "internal", tree.hiv.meta$source) #this is a workaround because otherwise the internal lines are removed when adding linetype as an aesthetic
  
geneotypeXregion.hiv.tree <- ggtree(tree.hiv, aes(linetype = source)) %<+% tree.hiv.meta +
  geom_tippoint(aes(color = region.CN, shape = region.CN, fill = hiv_subtype.F)) +
  scale_colour_manual(values = c("darkviolet", 000000, "black")) +
  scale_linetype_manual(values = c("solid", "dashed", "solid", "dotted")) +
  scale_shape_manual(values = c(22,21,24)) +
  #There is a bug for the interaction between fill and scale_shap_manual - this is a fix to make the legend work (however, important to doublecheck the scale is right because now it won't automatically match):
  guides(fill = guide_legend(override.aes=list(shape=c(21))),
         shape = guide_legend(override.aes=list(fill = "#7CAE00")))

geneotypeXregion.hiv.tree
ggsave(geneotypeXregion.hiv.tree, file = paste0(save_path, "geneotypeXregion.hiv.png"), width = 12, height = 30, units = "in")
```  

### Region - CN only
```{r plot region on SHARP and published tree, fig.height=60, fig.width=10}
region.CN.hiv.tree <- ggtree(tree.hiv) %<+% tree.hiv.meta +
  geom_tippoint(aes(colour = region.CN), stroke = 1.3) +
  geom_treescale() +
  scale_color_manual(values = c("purple4", "yellow2")) 

region.CN.hiv.tree
ggsave(region.CN.hiv.tree, file = paste0(save_path, "region.CN.hiv.tree.hiv.png"), width = 12, height = 30, units = "in")
``` 


### Risk
```{r plot region on SHARP and published tree, fig.height=60, fig.width=10}
risk.CN.hiv.tree <- ggtree(tree.hiv, aes(linetype = source)) %<+% tree.hiv.meta +
  scale_linetype_manual(values = c("solid", "dashed", "dotted")) +
  geom_tippoint(aes(colour = risk3.F), stroke = 1.3) +
  geom_treescale()

risk.hiv.tree
ggsave(risk.CN.hiv.tree, file = paste0(save_path, "risk.CN.hiv.tree.hiv.png"), width = 12, height = 30, units = "in")
``` 

